\Title

### 2020-10-29 manuscript/manuscript.tex
Start explaining iterative method on manuscript

\subsection{Equivalent sources implementation}

@@

>>>>

\subsection{Equivalent sources implementation}
\label{sec:eql-implementation}

@@

\section{Iteratively fitting source coefficients}

Reducing the number of sources helps, but it's not enough if we want to grid
a very large dataset.
Even if we can reduce the number of sources by one order of magnitude, having
too many data points would still increase the size of the Jacobian matrix until
the point of not being able to store it in memory.
This problem was already faced by other authors.

Here we propose an iterative algorithm for fitting the source coefficients
using a gradient boosting \citep{friedman2002} strategy and overlapping moving
windows.

Following equation \ref{eq:linear-problem} lets assume we have $N$ measurements
of a harmonic field on observations points $\{\mathbf{p}_i\}_{i=1}^{N}$ stored in
the vector $\mathbf{d} = (d_1, d_2, \ldots, d_N)^T$. We have defined a set of
$M$ point sources located on $\{\mathbf{q}_j\}_{j=1}^{M}$, whose coefficients
$\mathbf{c} = (c_1, c_2, \ldots, c_M)^T$ must be fitted so they can minimize
the misfit function defined on \ref{eq:misfit-unscaled}.

By defining a set of $K$ overlapping windows we can simplify the linear problem:

\begin{equation}
    \mathbf{d}
    = \mathbf{A} \mathbf{c}
    = \mathbf{A} \left( \sum_{k=1}^K \mathbf{c}_k \right)
    = \sum_{k=1}^K \mathbf{A} \mathbf{c}_k,
\end{equation}

\noindent where $\mathbf{c}_k$ are vectors full of zeros except those $m_k$
elements corresponding to the sources $\mathbf{q}_j$ that fall inside the
window $k$.

By applying gradient boosting we can fit each $\mathbf{c}_k$ vector
independently and iteratively by defining a residue vector $\mathbf{r}_k$ that
it's initialized with the observed data and updated on each iteration:

\begin{equation}
    \mathbf{r}_0 = \mathbf{d},
\end{equation}

\begin{equation}
    \mathbf{r}_k = \mathbf{r}_{k-1} - \mathbf{A}\mathbf{c}_k.
    \label{eq:residue}
\end{equation}

\noindent On each iteration the $\mathbf{c}_k$ vectors can be fitted so they
minimize the misfit function:

\begin{equation}
    \phi(\mathbf{c}_k) =
        \left\lVert
            \mathbf{d} - \mathbf{A}\mathbf{c}_k
        \right\rVert ^ 2
        + \lambda_d \left\lVert \mathbf{c}_k \right\rVert ^2,
\end{equation}

Because the $\mathbf{c}_k$ vector has only $m_k$ non-zero elements, the linear
system can be greatly simplified:

\begin{equation}
    \begin{split}
        \mathbf{A}\mathbf{c}_k = &
        \overbrace{
            \begin{bmatrix}
                a_{11} & a_{12} & a_{13} & \cdots & a_{1M} \\
                a_{21} & a_{22} & a_{23} & \cdots & a_{2M} \\
                a_{31} & a_{32} & a_{33} & \cdots & a_{3M} \\
                \vdots  & \vdots  & \vdots  & \cdots & \vdots \\
                a_{N1} & a_{N2} & a_{N3} & \cdots & a_{NM}
            \end{bmatrix}
        }^\text{$N \times M$ elements}
        \cdot
        \overbrace{
            \begin{bmatrix}
                0 \\
                {c_k}_2 \\
                0 \\
                \vdots  \\
                {c_k}_M \\
            \end{bmatrix}
        }^\text{$M$ elements}
        = \\ = &
        \underbrace{
            \begin{bmatrix}
                a_{12} & \cdots & a_{1M} \\
                a_{22} & \cdots & a_{2M} \\
                a_{32} & \cdots & a_{3M} \\
                \vdots  & \cdots & \vdots \\
                a_{N2} & \cdots & a_{NM}
            \end{bmatrix}
        }_\text{$N \times m_k$ elements}
        \cdot
        \underbrace{
            \begin{bmatrix}
                {c_k}_2 \\
                \vdots  \\
                {c_k}_M \\
            \end{bmatrix}
        }_\text{$m_k$ elements}
        = \tilde{\mathbf{A}} \tilde{\mathbf{c}}_k
        .
    \end{split}
\end{equation}

By removing the columns of the $\mathbf{A}$ matrix that are not needed in the
$k$-th iteration we greatly reduce the memory consumption to solve the linear
problem. Nevertheless, if the number of observation points ($N$) is too large,
the $\tilde{\mathbf{A}}$ matrix might be large enough for the available memory.

One more simplification can be carried out by ignoring the effect of the $m_k$
sources on the observation points that fall outside the $k$ window. This means
that the coefficients $\mathbf{c}_k$ will be fitted only taking into account
the data values on the $n_k$ observation points that fall inside the same
$k$-th window.
This allows to reduce the $\tilde{\mathbf{A}}$ matrix to a $\mathbf{B}$ matrix
with $n_k \times m_k$ elements:

\begin{equation}
    \mathbf{B} \tilde{\mathbf{c}}_k
    =
    \underbrace{
        \begin{bmatrix}
            a_{32} & \cdots & a_{3M} \\
            \vdots  & \cdots & \vdots \\
            a_{N2} & \cdots & a_{NM}
        \end{bmatrix}
    }_\text{$n_k \times m_k$ elements}
    \cdot
    \underbrace{
        \begin{bmatrix}
            {c_k}_2 \\
            \vdots  \\
            {c_k}_M \\
        \end{bmatrix}
    }_\text{$m_k$ elements}
    .
\end{equation}

These two simplifications help to greatly reduce the size of the linear
problem, reducing it to:

\begin{equation}
    \mathbf{d}_k = \mathbf{B} \tilde{\mathbf{c}}_k,
\end{equation}

\noindent where $\mathbf{d}_k$ is a vector containing only the measured data
values corresponding to the observation points that fall inside the $k$-th
window.
This linear problem can be solved using the same method described on Section
\ref{sec:eql-implementation}, obtaining values for the $\mathbf{c}_k$
coefficients and updating the residual through equation \ref{eq:residue}, i.e.
taking into account the effect of the fitted coefficients on every observation
point.
Worth noting that for updating the residue there's no need of computing the
entire Jacobian matrix.

After passing through every $K$ window, the residue $r_K$ should be
sufficiently small obtaining a set of source coefficients as the sum of the
$\mathbf{c}_k$ vectors:

\begin{equation}
    \mathbf{c} = \sum_{k=1}^K \mathbf{c}_k
\end{equation}

Shuffle the windows.

First we define a set of equivalent sources following one of the
source distributions described on the previous sections, e.g. the block-reduced
sources at variable depth. Secondly we define a set of overlapping moving
windows that cover the entire area of the data. Then we start the iterative
process: on each iteration we choose one window and select every data points
and sources that fall inside it, compute the Jacobian elements corresponding to
these subsets of data points and sources, fit the coefficients of the chosen
sources and finally remove the effect of these sources from the entire data,
obtaining a \emph{residue}. On the following iterations the same process is
carried out, but now the coefficients are fitted using the residue instead of
the original data.
